# SSDS 2019 Hands-on Session

## Workbook - Day 5

## Introduction to symbolic data

```{r}
library(dplyr)
library(magrittr)
```


*********************************************
### Assignment 1
*********************************************
Get a glimpse of the Iris dataset and display the first rows.
```{r}

```

*********************************************
### Assignment 2
*********************************************
Find minimal and maximal values of each feature in the Iris dataset for each species. This will allow us to create interval-valued symbolic variable as $\left[ x_{\min}, x_{\max} \right]$.
```{r}

```

```{r}

```


*********************************************
### Assignment 3
*********************************************
Import RSDA package and create a symbolic dataset from the Iris data. Use classic.to.sym function.
```{r}
library(RSDA)
```



## Univariate descriptive statistics

*********************************************
### Assignment 4
*********************************************
Calculate the symbolic sample mean for all variables in the Iris dataset. Use sym.mean function.
```{r}

```

*********************************************
### Assignment 5
*********************************************
Calculate the symbolic sample variance for all variables in the Iris dataset. Use sym.variance function.
```{r}

```



## Multivariate descriptive statistics

*********************************************
### Assignment 6
*********************************************
Calculate the empirical covariance between sepal length and sepal width variables of the Iris dataset. Use sym.cov function.
```{r}

```

*********************************************
### Assignment 7
*********************************************
Calculate the empirical correlation between Sepal length and Sepal width variables of the Iris dataset. Use sym.cor function.
```{r}

```



## Principal component analysis for symbolic data

*********************************************
### Assignment 8
*********************************************
Import blood pressure interval-valued data (lynne1) and plot the data using sym.scatterplot.

```{r}

```

*********************************************
### Assignment 9
*********************************************
Plot the blood pressure dataset using sym.scatterplot3d.

```{r}

```


*********************************************
### Assignment 10
*********************************************
Apply PCA vertices method to the blood pressure dataset. Use \texttt{sym.interval.pca} function.

```{r}

```



*********************************************
### Assignment 11
*********************************************
Apply PCA centers method to the blood pressure dataset. Use sym.interval.pca function.

```{r}

```






